---
title: "Assignment 1"
author: "Joseph Cerra"
date: "6/22/2020"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.path = "/Users/jcerra/Desktop/test/figures/")
```

#### Reading in the data from desktop
#### 1. Code for reading in the dataset and/or processing the data

```{r}
library(ggplot2)
activity <- read.csv("/Users/jcerra/Desktop/activity.csv")
str(activity)
summary(activity)
```

#### 2. Histogram of steps per day
```{r}
steps_day <- aggregate(steps ~ date, data = activity, FUN = sum, na.rm = TRUE)
hist(steps_day$steps, xlab = "Steps per Day", main = "Steps taken per day", col = "blue")
```

#### 3. Mean and Median num of steps per day

```{r}
meansteps <- mean(steps_day$steps)
mediansteps <- median(steps_day$steps)
meansteps
mediansteps
```

#### 4. Time series plot of the avg num of steps taken

```{r}
avg<- tapply(activity$steps, activity$interval, mean, na.rm=TRUE)
plot(names(avg), avg, xlab="5 minutes intervals", type="l", ylab="Average number of steps")
```

#### 5. The 5-minute interval that, on average, contains the maximum number of steps.

```{r}
maxavg<- max(avg)
maxinterval<- as.numeric(names(avg)[which(avg==max(avg))])
maxavg
maxinterval
```

#### 6. Inputting Missing data

```{r}
totalna <- sum(is.na(activity$steps))
imputedata <- activity
imputedata$steps[which(is.na(activity$steps))]<- as.vector(avg[as.character(activity[which(is.na(activity$steps)),3])])
#Create new data set that is a match to original but with imputed values
```

#### NAs are distributed equally over intervals however they occur on specific dates, therefore we can just take mean of the steps taken during these days.

#### 7. Histogram of the total number of steps taken each day after missing values are imputed

```{r}
stepseachday <- tapply(imputedata$steps, imputedata$date, sum, na.rm=TRUE)
qplot(stepseachday, xlab="No. of Steps Taken Each Day", ylab="Total Frequency", binwidth=500)
medianeach<- median(stepseachday)
meaneach<- mean(stepseachday)
medianeach
meaneach
```

#### Here we see that imputing missing values did not change the mean or median value

#### 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
imputedata$dayType<- ifelse(as.POSIXlt(imputedata$date)$wday %in% c(0,6), "weekends","weekdays")
aggregateData<- aggregate(steps ~ interval + dayType, data=imputedata, mean)
ggplot(aggregateData, aes(color = "", interval, steps)) + 
    geom_line() +
    facet_grid(dayType ~ .) +
    xlab("5 Minute Interval") + 
    ylab("Average number of steps")
```
